---
title: "Heart Failure Inpatient Analysis"
author: "Harrison Nguyen"
date: "25/06/2021"
output:  
  bookdown::html_document2:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Heart Failure Inpatient Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE, message=FALSE
)
knitr::opts_knit$set(
  root.dir = here::here()
)
```

```{r import}
library(magrittr)
library(targets)
library(tidyverse)
library(kableExtra)
library(ggplot2)
library(ggpubr)
library(acs)
source("R/definitions.R")
```

```{r data_load}


hf_journey <- targets::tar_read(hf_journey)
hf_journey_demo_prep <- targets::tar_read(hf_journey_demo_prep)

hf_discharge_letter_keys <- targets::tar_read(hf_discharge_letter_keys)

hf_journey_meds_admit_discharge <- targets::tar_read(hf_journey_meds_admit_discharge)
hf_meds_discharge_sum <- targets::tar_read(hf_meds_discharge_sum)


hf_echo_procedures <- targets::tar_read(hf_echo_procedures)
```

# Introduction

This document will examine the inpatient journeys of patients who were designated a heart failure related ICD-10 code. For instance, there will be a characterisation of the cohort in terms of their demographics, as well as the proportion that were discharged with specific medications and had undergone specific investigations and/or procedures.

# Cohort

The cohort under analysis are patient journeys that were designated the following ICD10 codes as a primary diagnosis:

```{r}
as.data.frame(hf_icd_codes()) %>% 
  knitr::kable()

```

```{r }
facility_list <- c("Royal North Shore")
cohort_journey_keys <- hf_journey %>%
  dplyr::filter(FACILITY %in%  facility_list, DIAGNOSIS_IMPORTANCE == "primary_diagnosis") %>%
  dplyr::pull(JOURNEY_KEY)



primary_journey <- hf_journey %>%
  dplyr::filter(JOURNEY_KEY %in% cohort_journey_keys )

primary_journey_demo <-hf_journey_demo_prep %>%
  dplyr::filter(JOURNEY_KEY %in% cohort_journey_keys )

primary_meds <-hf_journey_meds_admit_discharge %>%
  dplyr::filter(JOURNEY_KEY %in% cohort_journey_keys )

primary_meds_sum <- hf_meds_discharge_sum %>%
  dplyr::filter(JOURNEY_KEY %in% cohort_journey_keys )

cohort_encounter_keys <- primary_journey %>%
  dplyr::distinct(ENCNTR_KEY)

primary_echo <- hf_echo_procedures %>%
  dplyr::filter(ENCNTR_KEY %in% cohort_encounter_keys )
```

```{r}
n_journeys <- dplyr::n_distinct(primary_journey$JOURNEY_KEY)
n_encounters <- dplyr::n_distinct(primary_journey$ENCNTR_KEY)
n_patients <- dplyr::n_distinct(primary_journey$PERSON_KEY)
```
Furthermore, the cohort selected have an admission date between **`r as.Date(min(primary_journey_demo$JOURNEY_START))`** - **`r as.Date(max(primary_journey_demo$JOURNEY_START))`** and includes any journeys that includes an encounter at **`r paste(facility_list,sep=", ")`**.

The analysis therefore includes N = **`r n_journeys`** unique journeys with **`r n_encounters`** encounters for **`r n_patients`** unique patients.

```{r}

proportion_bar_plot <- function(df,column,title){
  df %>%
     ggplot2::ggplot(., aes(x=!!rlang::sym(column))) +
    ggplot2::geom_bar(aes(y = (..count..)/sum(..count..)*100),stat="count") +
    ggplot2::xlab(column) + 
    ggplot2::ylab("Percent (%)") + 
    ggtitle(paste0(title, " (N = ",nrow(df), ")")) +
    geom_text(
       aes(label=..count..,y= (..count..)/sum(..count..)*100),
       stat='count',
       vjust=-0.25
   )
}
```


# Demographics

Figure \@ref(fig:demo-plot) shows the distribution of ICD10 codes, age and sex.

```{r demo-plot, fig.cap="Demographics of journeys."}
icd10_plot<- proportion_bar_plot(primary_journey_demo,"DISPLAY_IDENTIFIER_CAP","ICD10 codes")

age_plot<- proportion_bar_plot(primary_journey_demo,"AGE_BINNED","AGE")

sex_plot<- proportion_bar_plot(primary_journey_demo,"GENDER","SEX")

ggpubr::ggarrange(icd10_plot, age_plot, sex_plot,
          labels = c("A", "B", "C"),
          ncol = 3, nrow = 1)
```
Figure \@ref(fig:code-cat-plot) shows the distribution of ICD10 codes conditioned on age and sex. The plots suggests that older age groups are more likely to be coded as having an ICD10 code of `I50.0 Heart failure` and are less likely to be coded as `I50.1 Left ventricular failure`. There seems to be no significant differences in diagnosis between sex.

```{r code-cat-plot, fig.cap="Distribution of ICD10 codes conditioned on sex and age."}


 age_plot<- primary_journey_demo %>%
  dplyr::count(AGE_BINNED,DISPLAY_IDENTIFIER_CAP) %>%
  dplyr::group_by(AGE_BINNED) %>%
  dplyr::mutate(prop = prop.table(n)) %>%
  ggplot2::ggplot(., aes(x=AGE_BINNED,y=prop,fill=DISPLAY_IDENTIFIER_CAP)) + 
  geom_bar(stat="identity",position=position_dodge(0.8),width=0.7) +
  scale_y_continuous(labels = scales::percent_format()) + 
  geom_text(aes(label= n,y=prop),position=position_dodge(0.8),vjust=-0.5) +
  labs(y="Percent (%)") +
  scale_fill_discrete(name = "ICD10 code")

 sex_plot<- primary_journey_demo %>%
  dplyr::count(GENDER,DISPLAY_IDENTIFIER_CAP) %>%
  dplyr::group_by(GENDER) %>%
  dplyr::mutate(prop = prop.table(n)) %>%
  ggplot2::ggplot(., aes(x=GENDER,y=prop,fill=DISPLAY_IDENTIFIER_CAP)) + 
  geom_bar(stat="identity",position=position_dodge(0.8),width=0.7) +
  scale_y_continuous(labels = scales::percent_format()) + 
  geom_text(aes(label= n,y=prop),position=position_dodge(0.8),vjust=-0.5) +
  labs(y="Percent (%)") +
  scale_fill_discrete(name = "ICD10 code")
 
  

ggpubr::ggarrange(age_plot, sex_plot,
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```


```{r}
postcode_plot<- primary_journey_demo %>%
   ggplot2::ggplot(., aes(x=CURRENT_POSTCODE)) +
  ggplot2::geom_bar() +
  ggplot2::xlab("POSTCODE") + ggplot2::ylab("COUNT") + ggtitle(paste0("POSTCODE (N = ",n_journeys, ")")) +
  geom_text(
     aes(label=..count..),
     stat='count',
     nudge_y=0.5,
     va='bottom'
 )
postcode_plot
```

# Journey Characteristics

Figure \@ref(fig:journey-char-plot) shows the characteristics of the journeys such as the presentation mode, whether the journey involved a transfer and the separation mode. Those presented as `Ambulance Triage` are defined as journeys that hadtheir jorueny first facility as `Royal North Shore` and whether their triage form contained the presence of `atami` or `etami` or `cath lab`.

Most encounters are admitted into ED and most journeys either arrived by ambulance or as a self presentation. 

```{r journey-char-plot, fig.cap="Characteristics of journeys."}

p1<-proportion_bar_plot(primary_journey, "ENCLASS", "Encounter_type")

p2<- proportion_bar_plot(primary_journey_demo, "PRESENTATION_MODE", "PRESENTATION_MODE")
p3<-proportion_bar_plot(primary_journey_demo, "TRANSFER_STATUS", "TRANSFER_STATUS")
p4<-proportion_bar_plot(primary_journey_demo, "SEPARATION_MODE", "SEPARATION_MODE")

ggpubr::ggarrange(p1, p2, p3,p4,
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 2)

```

# Discharge Medications


Discharge medications were defined as journeys that had a discharge letter and excluded journeys that had a separation mode of either **`r paste(exclusion_criteria_meds(),sep=",")`**.

```{r}
hf_discharge_journeys <- primary_journey %>%
  dplyr::filter(ENCNTR_KEY %in% hf_discharge_letter_keys)

n_journeys_w_discharge <- dplyr::n_distinct(hf_discharge_journeys$JOURNEY_KEY)

## need to consider eplerenone
## https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4814010/
```

**`r round(n_journeys_w_discharge/n_journeys*100,1)`**% (**`r n_journeys_w_discharge`**/**`r n_journeys`**) had a discharge letter and of those,**`r round(nrow(primary_meds_sum)/n_journeys_w_discharge*100,1)`**% (**`r  nrow(primary_meds_sum)`**/**`r n_journeys_w_discharge`**) remained after the exclusion criteria.



```{r}
ace<- hf_journey_meds_admit_discharge %>%
  dplyr::filter(MEDICATION == "A2RB/ACE Inhibitor") %>%
  dplyr::distinct(MEDICATION_NAME)

beta<- hf_journey_meds_admit_discharge %>%
  dplyr::filter(MEDICATION == "Beta blocker") %>%
  dplyr::distinct(MEDICATION_NAME)

diuretic<- hf_journey_meds_admit_discharge %>%
  dplyr::filter(MEDICATION == "Other diuretic") %>%
  dplyr::distinct(MEDICATION_NAME)
```

The medications examined include:

* ACE inhibitors such as **`r paste0(ace$MEDICATION_NAME[1:3],collapse=", ")`**,
* beta blockers such as **`r paste0(beta$MEDICATION_NAME[1:3],collapse=", ")`** and
* diuretics such as **`r paste0(diuretic$MEDICATION_NAME[1:3],collapse=", ")`**.

Figure \@ref(fig:meds-plot) shows the proportion of journeys that were discharged with the relevant medication.


```{r meds-plot, fig.cap="Proportion of journeys with discharge medication."}
primary_meds_sum %>%
  gather(medication, val, `Other diuretic`:`Beta blocker`) %>%
  mutate(val = val > 0) %>%
  group_by(medication) %>%
  summarise(percent = sum(val)/n()*100,n = sum(val), .groups = "drop") %>%
  mutate(medication = fct_reorder(medication, percent)) %>%
  ggplot2::ggplot(., aes(x=medication,y=percent))+
  ggplot2::geom_bar(stat="identity") +
  geom_text(
     aes(label=n),
     nudge_y=1.5,
     va='bottom'
 ) +
  labs(x="Medication",y="Percent (%)",
       title = paste0("Medication (N = ", nrow(primary_meds_sum), ")"))
  
```

# Investigations

We examine the number of echocardiograms that were performed in the cohort. Currently, the extract only contains transoesophageal echocardiographies or echocardiograms that were performed via an endoscope. 

There are **`r nrow(hf_echo_procedures)`** transoesophageal echocardiograms in the extract. However **`r nrow(primary_echo)`** of the echos were associated with the cohort of interest.

Figure \@ref(fig:echo-time-delta-plot) shows the distribution of the admission time and the order of the echo procedure.

```{r echo-time-delta-plot, fig.cap="Distribution of the time difference between admission time and echocardiogram."}
ggplot(hf_echo_procedures, aes(x=PROCEDURE_ADMIT_TIME_DELTA)) + geom_histogram(bins=10) +
  labs(x="PROCEDURE_ADMIT_TIME_DELTA (hours)")
```


# Discharge Diagnosis and ICD10 Diagnosis Concordance
```{r}
hf_snomed_encounters <- targets::tar_read(hf_snomed_encounters)
hf_icd10_encounters <- targets::tar_read(hf_icd10_encounters)
fac_snomed_encounter <- hf_snomed_encounters %>%
  dplyr::filter(FACILITY %in% facility_list)

fac_icd10_encounter <- hf_icd10_encounters %>%
  dplyr::filter(FACILITY %in% facility_list)

```

We examine encounters with either a ICD10 or discharge (SNOMED) diagnosis of heart failure over the same period of time and facilities. This results in **`r nrow(fac_icd10_encounter)`** ICD10 encounter and **`r nrow(fac_snomed_encounter)`** SNOMED encounters with the following list of SNOMED codes:

```{r}
as.tibble(unique(fac_snomed_encounter$SOURCE_STRING_CAP)) %>% knitr::kable()
```

```{r}
comb_encounters<-fac_icd10_encounter %>%
  dplyr::select(ENCNTR_KEY, SOURCE_STRING_CAP) %>%
  dplyr::full_join(
    dplyr::select(fac_snomed_encounter,ENCNTR_KEY,SOURCE_STRING_CAP),
    by="ENCNTR_KEY",
    suffix=c("_ICD10","_SNOMED")
  ) %>%
  dplyr::mutate(HAS_ICD10 = !is.na(SOURCE_STRING_CAP_ICD10),
                HAS_SNOMED = !is.na(SOURCE_STRING_CAP_SNOMED))

snomed_icd10<- table(comb_encounters$HAS_ICD10,comb_encounters$HAS_SNOMED) 

prop_snomed_icd10<- prop.table(table(comb_encounters$HAS_ICD10,comb_encounters$HAS_SNOMED)) 

```

Taking the union of the two types of diagnosis codes together, there are **`r nrow(comb_encounters)`** unique encounters.

Table \@ref(tab:icd10-tab) shows a contingency table between discharge diagnosis and ICD10 codes where only **`r round(prop_snomed_icd10[2,2]*100,1)`**% have both an ICD10 and SNOMED heart failure code.

```{r icd10-tab}


snomed_icd10 %>% 
  knitr::kable(digits=2,caption = "Contigency table of discharge diagnosis and ICD10 code.") %>%
  kableExtra::add_header_above(c("HAS_ICD10"=1,"HAS_SNOMED"=2))

prop_snomed_icd10 %>% 
  knitr::kable(digits=2, caption = "Contigency table of discharge diagnosis and ICD10 code as percentage.") %>%
  kableExtra::add_header_above(c("HAS_ICD10"=1,"HAS_SNOMED"=2))
```

# Secondary Diagnosis

The above results focused on those with a primary ICD10 heart failure diagnosis. The next section examines those with a `secondary` or `additional` ICD10 diagnosis.

```{r}
cohort_journey_keys <- hf_journey %>%
  dplyr::filter(FACILITY %in%  facility_list, DIAGNOSIS_IMPORTANCE == "secondary_diagnosis") %>%
  dplyr::pull(JOURNEY_KEY)

secondary_journey_demo <- hf_journey_demo_prep %>%
  dplyr::filter(JOURNEY_KEY %in% cohort_journey_keys )

icd10_plot<- proportion_bar_plot(secondary_journey_demo,"DISPLAY_IDENTIFIER_CAP","Additional Diagnosis")

icd10_plot
```
The cohort for a secondary diagnosis of heart failure is larger (N = **`r nrow(secondary_journey_demo)`**) than those diagnosed with a primary diagnosis (N = **`r nrow(primary_journey_demo)`**) and also includes encounters coded with `chronic heart failure`.
