---
title: "Validating ACS rules"
author: "Harrison"
date: "`r Sys.Date()`"
output:  
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE, message=FALSE
)
```

```{r import}
library(acs)
library(magrittr)
library(dplyr)
library(knitr)
library(kableExtra)
```

```{r}
load("../data/journey_analysis_base.rda")

```

```{r}
n_encounters <- dplyr::n_distinct(journey_analysis_base$encntr_key)

n_journeys <- dplyr::n_distinct(journey_analysis_base$JOURNEY_KEY)
```
# Introduction

We aim to investigate the accuracy of identifying STEMI patients based on various ACS rules and troponin results.
The cohort under investigation are encounters across the NSLHD between the dates **`r as.Date(min(journey_analysis_base$admit_dttm))`** - **`r as.Date(max(journey_analysis_base$admit_dttm))`**. 

Across this period there are **`r n_encounters`** unique encounters and **`r n_journeys`** unique journeys (a collection of encounters that have occurred within a specified time interval).


```{r}
stemi <-journey_analysis_base %>%
  dplyr::mutate(STEMI = stringr::str_detect(icd10_diagnosis_list,"[^N]STEMI") | stringr::str_starts(icd10_diagnosis_list,"STEMI")) %>%
  dplyr::select(encntr_key,journey_acs_rule1,journey_acs_rule2,journey_above_normal_trop,JOURNEY_KEY,STEMI) %>%
  dplyr::distinct(JOURNEY_KEY,.keep_all=TRUE)
```

# Troponin Results

There are three different rules that can define a ACS patient

## ACS Rule 1: 
 **Relative to first troponin reading of journey:**
 Find the first Troponin result for the journey 
 (According to `result_start_dttm`.) 
 
 Now, for every subsequent troponin result, if a given result's value
 (`result_discrete_value`) exceeds the first by `> %30`, ie: 
 
 `result[i] > 1.3*result[1]` for *i* within *2* and *n* where *n* is 
 the total number of results for that journey. 
 
  For a result that exceeds this, set `pct_change_thresh_flag <- TRUE`,
 
 If this flag is set for any result row in a particular journey, 
 for all rows in that journey set `group_pass_pct_thresh <- TRUE`

```{r}
kable_confusion_matrix <- function(table, row_name, col_name){
  span <- c(1,2)
  names(span) <- c(" ", col_name)
  result <- knitr::kable(table) %>%
  kableExtra::add_header_above(span) %>%
  kableExtra::pack_rows(row_name, 1, 2) %>%
  kableExtra::kable_styling(full_width = F)
  result
}

```
 
```{r}
kable_confusion_matrix(table(stemi$journey_acs_rule1,stemi$STEMI), "ACS rule1", "ICD-10 STEMI")
```

## ACS Rule 2: 
 **Relative to absolute value:**
 
 This rule is simpler than **ACS Rule 1**, and simply involves comparing
 each `result_discrete_value` with a large number, chosen to be `1000`.
 
 If a result exceeds that, flag its row with `abs_high_thresh_flag <- TRUE`,

 and flag all rows in that journey with `group_pass_abs_high_thresh <- TRUE`
 
```{r}
kable_confusion_matrix(table(stemi$journey_acs_rule2,stemi$STEMI), "ACS rule 2", "ICD-10 STEMI")
```


## Troponin Rule 3:
  **Compare** `result_discrete_value` **to value of** `normal_high` 
  *(Upper threshhold for normal).*
  
  For each row, a column, `normal_high`, is given, specifying what the upper
  limit for that result **SHOULD** be for it to be considered "normal". 
  
  For each row in a journey, if `result_discrete_value > normal_high`,
 flag that row with `above_normal_flag <- TRUE` and for each row in the 
  journey, `group_above_normal <- TRUE`.
 
```{r}
kable_confusion_matrix(table(stemi$journey_above_normal_trop,stemi$STEMI),"HIGH TROPONIN", "ICD-10 STEMI")
```

