---
title: "Validating ACS rules"
author: "Harrison"
date: "`r Sys.Date()`"
output:  
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE, message=FALSE
)
```

```{r import}
library(acs)
library(magrittr)
library(dplyr)
library(knitr)
library(kableExtra)
```

```{r}
#load("../data/journey_analysis_base.rda")
journey_analysis_base <-read.csv(here::here("data-raw", "journey_analysis_base.csv"), stringsAsFactors = FALSE)
#data("journey_analysis_base")
```

```{r}
n_encounters <- dplyr::n_distinct(journey_analysis_base$encntr_key)

n_journeys <- dplyr::n_distinct(journey_analysis_base$JOURNEY_KEY)
```
# Introduction

We aim to investigate the accuracy of identifying STEMI patients based on various ACS rules and troponin results.
The cohort under investigation are encounters across the NSLHD between the dates **`r as.Date(min(journey_analysis_base$admit_dttm))`** - **`r as.Date(max(journey_analysis_base$admit_dttm))`**. 

Across this period there are **`r n_encounters`** unique encounters and **`r n_journeys`** unique journeys (a collection of encounters that have occurred within a specified time interval).


```{r}
stemi <-journey_analysis_base %>%
  dplyr::filter(admit_dttm > as.POSIXct("2016-05-01") &
                  admit_dttm < as.POSIXct("2018-04-30")) %>%
  dplyr::mutate(STEMI = stringr::str_detect(icd10_diagnosis_list,"[^N]STEMI") | stringr::str_starts(icd10_diagnosis_list,"STEMI")) %>%
  dplyr::distinct(JOURNEY_KEY,.keep_all=TRUE)
```

# Troponin Results

There are three different rules that can define a ACS patient

## ACS Rule 1
 **Relative to first troponin reading of journey:**
 Find the first Troponin result for the journey 
 (According to `result_start_dttm`.) 
 
 Now, for every subsequent troponin result, if a given result's value
 (`result_discrete_value`) exceeds the first by `> %30`, ie: 
 
 `result[i] > 1.3*result[1]` for *i* within *2* and *n* where *n* is 
 the total number of results for that journey. 

```{r}
kable_confusion_matrix <- function(table, row_name, col_name){
  span <- c(1,2)
  names(span) <- c(" ", col_name)
  result <- knitr::kable(table) %>%
  kableExtra::add_header_above(span) %>%
  kableExtra::pack_rows(row_name, 1, 2) %>%
  kableExtra::kable_styling(full_width = F)
  result
}

```
 
```{r}
kable_confusion_matrix(table(stemi$journey_acs_rule1,stemi$STEMI), "ACS rule1", "ICD-10 STEMI")
```

## ACS Rule 2
 **Relative to absolute value:**
 
 This rule is simpler than **ACS Rule 1**, and simply involves comparing
 each `result_discrete_value` with a large number, chosen to be `1000`.
 
```{r}
kable_confusion_matrix(table(stemi$journey_acs_rule2,stemi$STEMI), "ACS rule 2", "ICD-10 STEMI")
```


## Troponin Rule 3
  **Compare** `result_discrete_value` **to value of** `normal_high` 
  *(Upper threshhold for normal).*
  
  For each row, a column, `normal_high`, is given, specifying what the upper
  limit for that result **SHOULD** be for it to be considered "normal". 
  
```{r}
kable_confusion_matrix(table(stemi$journey_above_normal_trop,stemi$STEMI),"HIGH TROPONIN", "ICD-10 STEMI")
```
### Other Criteria

#### Original 3 Criteria
We also look at 3 other inclusions criteria being:

1. The “Reason for Visit” (free text field) for the presentation contained any of the ACS-related symptoms or keywords described in Supplementary Information Part 1.
2. Orders were placed in the EMR for a troponin test OR a 12 lead ECG OR for any of the following investigations: coronary angiogram, exercise stress test, stress echocardiogram, sestamibi scan, CT coronary angiogram, CT pulmonary angiogram.
3.  The encounter contained a scanned 12-lead ECG image.

```{r}
cohort_flags <- stemi %>%
  dplyr::rename_all(toupper) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    CF1_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[1]),
    CF2_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[2]),
    CF3_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[3]),
    CF5_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[5]),
    CF6_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[6]),
    CF7_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[7])
  ) %>%
  dplyr::mutate(
    CRITERIA = CF1_derived | CF3_derived | CF7_derived
  )
```

```{r}
criteria_table <- table(cohort_flags$CRITERIA,cohort_flags$STEMI)
kable_confusion_matrix(criteria_table,"SELECTION CRITERIA", "ICD-10 STEMI")
```

#### Alternate Criteria
Given that the above results show a lot of false positives, we strengthen the second criteria and only include patients with **ONLY** troponin orders. In other words, the new criteria is:

1. The “Reason for Visit” (free text field) for the presentation contained any of the ACS-related symptoms or keywords described in Supplementary Information Part 1.
2. **Orders were placed in the EMR for a troponin test.**
3.  The encounter contained a scanned 12-lead ECG image.

```{r}
load(file.path(here::here("data"),"troponin_encounter_order.rda"))
```

```{r}
troponin_order_flag <- troponin_encounter_order %>% 
  dplyr::distinct(ENCNTR_KEY,.keep_all=TRUE)%>%
  dplyr::mutate(TROPONIN_FLAG=TRUE) %>%
  dplyr::select(ENCNTR_KEY,TROPONIN_FLAG)


new_cohort_flags <- cohort_flags %>%
  dplyr::left_join(troponin_order_flag, by="ENCNTR_KEY") %>%
  dplyr::mutate(
    TROPONIN_FLAG = tidyr::replace_na(TROPONIN_FLAG,FALSE)
      )%>%
  dplyr::mutate(
    CRITERIA_NEW = CF1_derived | TROPONIN_FLAG | CF7_derived
  )
```

```{r}
criteria_new_table <- table(new_cohort_flags$CRITERIA_NEW,new_cohort_flags$STEMI)
kable_confusion_matrix(criteria_new_table,"SELECTION CRITERIA", "ICD-10 STEMI")
```
```{r}
new_fp <- criteria_new_table[2][1]

old_fp <- criteria_table[2][1]

diff <- old_fp - new_fp
```

The change of the second criteria of only using the troponin orders has not made any significant difference, the number of false positives has decreased by `r diff` out of the original `r old_fp`, a decrease of `r diff/old_fp*100`%.
