---
title: "Validating ACS rules"
author: "Harrison"
date: 19/4/2021
output:  
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE, message=FALSE
)
```

```{r import}
library(acs)
library(magrittr)
library(dplyr)
library(knitr)
library(kableExtra)
```

```{r}
#load("../data/journey_analysis_base.rda")
journey_analysis_base <-read.csv(here::here("data-raw", "journey_analysis_base.csv"), stringsAsFactors = FALSE)
#data("journey_analysis_base")

STEMI_SNOMED <- "(?i)acute( inferior | anterior |\\s)st segment elevation|(ST elevation myocardial infarction)|(Acute STEMI)"


load("../data/diagnosis_data.rda")
```

```{r}
snomed <-  diagnosis_data %>%
  dplyr::filter(confirmation_status == "Confirmed" & (classification == "ED Medical" | classification == "Medical" | classification == "ED Nursing")) %>%
  dplyr::filter(n_source_vocabulary_disp == "SNOMED CT")

snomed_stemi <- snomed %>%
  dplyr::filter(stringr::str_detect(source_string_cap, STEMI_SNOMED))
```

```{r}

# should be standard
`%notin%` <- Negate(`%in%`)

NSLHD <- c("Royal North Shore", "Hornsby", "Ryde", "Manly", "Mona Vale")
CCLHD <- c("Gosford", "Woy Woy", "Wyong", "Long Jetty")

is.ACS <- function(.x) {
  any(.x %in% c("STEMI", "NSTEMI", "UNSTABLE_ANGINA", "UNSPECIFIED_ACS"), 
      na.rm=T)
}

internal_transfers = c(
  # "Internal Ambulance / Transport",
  # "Retrieval",
  "Internal Bed / Wheelchair"
)

journey_analysis_base %<>%  
  dplyr::group_by(JOURNEY_KEY) %>%
  dplyr::filter(
    any(enclass == "ED.Admitted" & facility %in% NSLHD & admit_mode %notin% internal_transfers, na.rm=T))%>%
  dplyr::ungroup()
```

```{r}
snomed_journeys <- snomed_stemi %>%
  dplyr::select(encntr_key,source_string_cap) %>%
  dplyr::right_join(journey_analysis_base, by="encntr_key") %>%
  dplyr::filter(!is.na(source_string_cap))
```

```{r}
stemi <-journey_analysis_base %>%
  dplyr::filter(admit_dttm > as.POSIXct("2016-05-01") &
                  admit_dttm < as.POSIXct("2018-04-30")) %>%
  dplyr::mutate(STEMI = stringr::str_detect(icd10_diagnosis_list,"[^N]STEMI") | stringr::str_starts(icd10_diagnosis_list,"STEMI")) %>%
  dplyr::distinct(JOURNEY_KEY,.keep_all=TRUE) %>%
  dplyr::mutate(SNOMED_STEMI = JOURNEY_KEY %in% snomed_journeys$JOURNEY_KEY)
```

```{r}
n_encounters <- dplyr::n_distinct(stemi$encntr_key)

n_journeys <- dplyr::n_distinct(stemi$JOURNEY_KEY)

n_icd10_stemi <- sum(stemi$STEMI)

n_snomed_stemi <- sum(stemi$SNOMED_STEMI)
```

# Introduction

## Aim    
We aim to investigate the accuracy of identifying STEMI patients based on various ACS rules and troponin results. We seek to analyse the amount of STEMI patients and non-STEMI patients that could be included in a cohort for a future extract if these different selection criteria is applied. 


## Cohort
The cohort under investigation are journeys with admission dates between **`r as.Date(min(journey_analysis_base$admit_dttm))`** - **`r as.Date(max(journey_analysis_base$admit_dttm))`**. 

The cohort is further filtered for those who are ED admitted in NSLHD and not an internal transfer    .

Across this period there are **`r n_encounters`** unique encounters and **`r n_journeys`** unique journeys (a collection of encounters that have occurred within a specified time interval).

The ICD-10 STEMI definitions can be found in an earlier analysis, shown [here](https://pages.github.sydney.edu.au/speed-extract/acs/articles/snomed_icd10_stemi.html#icd-10-1) and the SNOMED-STEMI definitions can be found at this [link](https://pages.github.sydney.edu.au/speed-extract/acs/articles/snomed_icd10_stemi.html#snomed-code-1).


All percentages have used the number of unique journeys, **`r n_journeys`**, as the denominator. In total there are

- **`r n_icd10_stemi`** ICD-10 STEMIS and 
- **`r n_snomed_stemi`** SNOMED STEMIS.

# Troponin Results

There are three different rules that can define a ACS patient. The troponin selection rule based on an above normal value given by the assay is not included as it is deemed not relevant.

## Comparison against Benchmark (ACS Rule 1)
 **Relative to first troponin reading of journey:**
 Find the first Troponin result for the journey 
 (According to `result_start_dttm`.) 
 
 Now, for every subsequent troponin result, if a given result's value
 (`result_discrete_value`) exceeds the first by `> %30`, ie: 
 
 `result[i] > 1.3*result[1]` for *i* within *2* and *n* where *n* is 
 the total number of results for that journey. 

```{r}
kable_confusion_matrix <- function(table, row_name, col_name){
  span <- c(1,2)
  names(span) <- c(" ", col_name)
  result <- knitr::kable(table) %>%
  kableExtra::add_header_above(span) %>%
  kableExtra::pack_rows(row_name, 1, 2) %>%
  kableExtra::kable_styling(full_width = F)
  result
}

table_list <- list()
i <- 1
criteria_name <- list()
```
 
The table shows a confusion matrix between those who have either been defined as a ICD10-STEMI or not and whether they have satisfied the troponin criteria.  
 
```{r}
stemi_table_name <-  "ICD-10 STEMI"
criteria_name[[i]] <- 'ACS rule1'

table_list[[i]] <- table(stemi$journey_acs_rule1,stemi$STEMI)
kable_confusion_matrix(table_list[[i]], criteria_name[[i]],stemi_table_name)
i <- i + 1
```

## High Threshold (ACS Rule 2)
 **Relative to absolute value:**
 
 This rule is simpler than **ACS Rule 1**, and simply involves comparing
 each `result_discrete_value` with a large number, chosen to be `1000`.
 
```{r}
criteria_name[[i]] <- 'ACS rule2'
table_list[[i]]  <- table(stemi$journey_acs_rule2,stemi$STEMI )
kable_confusion_matrix(table_list[[i]], criteria_name[[i]], stemi_table_name)

i <- i + 1
```

<!-- ## ACS Rule 3 -->
<!--   **Compare** `result_discrete_value` **to value of** `normal_high`  -->
<!--   *(Upper threshhold for normal).* -->

<!--   For each row, a column, `normal_high`, is given, specifying what the upper -->
<!--   limit for that result **SHOULD** be for it to be considered "normal".  -->

<!-- ```{r} -->

<!-- criteria_name[[i]] <- 'ACS Rule 3' -->
<!-- table_list[[i]] <- table(stemi$journey_above_normal_trop,stemi$STEMI) -->
<!-- kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name) -->

<!-- i <- i + 1 -->
<!-- ``` -->

## Combining ACS rules

We analyse the results in a similar manner by combining the troponin rules.

<!-- Note that   the threshold used for Rule 2 is generally higher than the above normal range used in Rule 3, thus the cohort of patients who satisfy rule 2 is a subset of rule 3. -->

### ACS rule 1 OR ACS rule 2

```{r}
criteria_name[[i]] <- 'ACS 1 Or 2'
table_list[[i]] <- table(stemi$journey_acs_rule1 | stemi$journey_acs_rule2,stemi$STEMI)
kable_confusion_matrix(table_list[[i]],criteria_name[[i]] , stemi_table_name)

i <- i + 1
```

<!-- ### ACS rule 1 OR ACS rule 3 -->

<!-- ```{r} -->
<!-- criteria_name[[i]] <- 'ACS 1 Or 3' -->

<!-- table_list[[i]] <- table(stemi$journey_acs_rule1 | stemi$journey_above_normal_trop,stemi$STEMI) -->
<!-- kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name) -->

<!-- i <- i + 1 -->
<!-- ``` -->


# Other Criteria

## Original 3 Criteria
We also look at 3 other inclusions criteria being:

1. The “Reason for Visit” (free text field) for the presentation contained any of the ACS-related symptoms or keywords described in Supplementary Information Part 1.
2. Orders were placed in the EMR for a troponin test OR a 12 lead ECG OR for any of the following investigations: coronary angiogram, exercise stress test, stress echocardiogram, sestamibi scan, CT coronary angiogram, CT pulmonary angiogram.
3.  The encounter contained a scanned 12-lead ECG image.

```{r}
cohort_flags <- stemi %>%
  dplyr::rename_all(toupper) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    CF1_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[1]),
    CF2_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[2]),
    CF3_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[3]),
    CF5_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[5]),
    CF6_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[6]),
    CF7_derived = as.integer(intToBits(COHORT_FLAG_DERIVED)[7])
  ) %>%
  dplyr::mutate(
    CRITERIA = CF1_derived | CF3_derived | CF7_derived
  )
```

```{r}
criteria_name[[i]] <- '3 Original Criteria'
table_list[[i]] <- table(cohort_flags$CRITERIA,cohort_flags$STEMI)
kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name)

i <- i+ 1
```

## Alternate Criteria
Given that the above results show a lot of false positives, we strengthen the second criteria and only include patients with **ONLY** troponin orders. In other words, the new criteria is:

1. The “Reason for Visit” (free text field) for the presentation contained any of the ACS-related symptoms or keywords described in Supplementary Information Part 1.
2. **Orders were placed in the EMR for a troponin test.**
3.  The encounter contained a scanned 12-lead ECG image.

```{r}
load(file.path(here::here("data"),"troponin_encounter_order.rda"))
```

```{r}
troponin_order_flag <- troponin_encounter_order %>% 
  dplyr::distinct(ENCNTR_KEY,.keep_all=TRUE)%>%
  dplyr::mutate(TROPONIN_FLAG=TRUE) %>%
  dplyr::select(ENCNTR_KEY,TROPONIN_FLAG)


new_cohort_flags <- cohort_flags %>%
  dplyr::left_join(troponin_order_flag, by="ENCNTR_KEY") %>%
  dplyr::mutate(
    TROPONIN_FLAG = tidyr::replace_na(TROPONIN_FLAG,FALSE)
      )%>%
  dplyr::mutate(
    CRITERIA_NEW = CF1_derived | TROPONIN_FLAG | CF7_derived
  )
```

```{r}
criteria_name[[i]] <- 'Adjusted 3 Criteria'
table_list[[i]] <- table(new_cohort_flags$CRITERIA_NEW,new_cohort_flags$STEMI)
kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name)

i<- i + 1
```
```{r}
new_fp <- table_list[[i-1]][2][1]

old_fp <- table_list[[i-2]][2][1]

diff <- old_fp - new_fp
```

The change of the second criteria of only using the troponin orders has not made any significant difference, the number of false positives has decreased by `r diff` out of the original `r old_fp`, a decrease of `r diff/old_fp*100`%.

## Analysing the Permutations of the Criteria

### Presentation of ACS-related symptoms (Criteria 1)
```{r}
criteria_name[[i]] <- 'Only C1'
table_list[[i]] <- table(as.logical(new_cohort_flags$CF1_derived),new_cohort_flags$STEMI)
kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name)

i<- i + 1
```

### Order of Troponin test (Adjusted Criteria 2)
```{r}
criteria_name[[i]] <- 'Only  Adj. C2'
table_list[[i]] <- table(as.logical(new_cohort_flags$TROPONIN_FLAG),new_cohort_flags$STEMI)
kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name)

i <- i + 1
```

### Has 12-lead ECG image (Criteria 3)
```{r}
criteria_name[[i]] <- 'Only C3'
table_list[[i]] <- table(as.logical(new_cohort_flags$CF7_derived),new_cohort_flags$STEMI)
kable_confusion_matrix(table_list[[i]] ,criteria_name[[i]], stemi_table_name)

i <- i + 1
```


### Criteria 1 OR 2

```{r}
criteria_name[[i]] <- 'C1 Or Adj. C2'
table_list[[i]] <- table(as.logical(new_cohort_flags$CF1_derived) | as.logical(new_cohort_flags$TROPONIN_FLAG),new_cohort_flags$STEMI)
kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name)

i<- i + 1
```

### Criteria 1 OR 3

```{r}
criteria_name[[i]] <- 'C1 Or C3'
table_list[[i]] <- table(as.logical(new_cohort_flags$CF1_derived) | as.logical(new_cohort_flags$CF7_derived),new_cohort_flags$STEMI)
kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name)

i<- i + 1
```

### Criteria 2 OR 3

```{r}
criteria_name[[i]] <- 'Adj. C2 Or C3'
table_list[[i]] <- table(as.logical(new_cohort_flags$TROPONIN_FLAG) | as.logical(new_cohort_flags$CF7_derived),new_cohort_flags$STEMI)
kable_confusion_matrix(table_list[[i]],criteria_name[[i]], stemi_table_name)

i<- i + 1
```

# Summary
  
We summarise the various cohort selection criteria using *precision* and *sensitivity* and combine these metrics through the F1 score, which is the harmonic mean of precision and sensitivity. 


```{r}
precision <- function(table){
  tp <- table[4]
  fp <- table[2]
  
  return(tp/(tp+fp))
}

specificity <- function(table){
  tn <- table[1]
  fp <- table[3]
  return(tn/(tn+fp))
}

sensitivity <- function(table){
  tp <- table[4]
  fn <- table[3]
  
  return(tp/(tp+fn))
}
```

```{r}
df <- data.frame(
                 precision = double(),
                 sensitivity = double())
for(i in 1:length(table_list)){
  df %<>% dplyr::add_row(
     precision = precision(table_list[[i]]),
      sensitivity= sensitivity(table_list[[i]])
  )
}

df %<>% dplyr::mutate(f_score = 2*precision*sensitivity/(precision + sensitivity))

rownames(df) <- unlist(criteria_name, use.names=FALSE)

knitr::kable(df %>% dplyr::mutate_if(is.numeric, format, digits=2), align="r") %>%
  kableExtra::kable_styling()%>%
  kableExtra::row_spec(.,c(2,7,6,9), bold= TRUE)
  
```

Using Rule 2, choosing the cohort by using a threshold on the troponin score, produces the highest precision however is the lowest sensitivity out of all the options. In other words, the high troponin score selects a very small cohort where **`r table_list[[2]][2,2]/sum(table_list[[2]][2,])*100`**\% are ICD-10-STEMI but misses **`r (1-df['ACS rule2','sensitivity'])*100`**\% STEMI's.

There little difference between including orders for procedures such as angiograms/ECGs or not as a selection criteria. These selection methods produce the highest sensitivity, at the same time, produces the highest number of false positives (large number of non-STEMI's) and thus lower precision when compared to using the troponin test results.

Furthermore, including a cohort with a troponin order (*Only Adj C2*) produces a relatively high sensitivity and precision when compared with the other criteria individually. In particular, it produces a precision that is **`r df['Only  Adj. C2','precision']/df['Only C3','precision']`** times higher than using only *C3*. Combining *C1* with *Adj. C2* is beneficial, as it improves the sensitivity without sacrificing a lot of precision.


# SNOMED

The above analysis is repeated looking at SNOMED codes as well. Only the first table is shown but all results are summarised with precision and sensitivity.

```{r}
table_list_snomed <- list()
i <- 1
stemi_table_name <- "SNOMED STEMI"

table_list_snomed[[i]] <- table(stemi$journey_acs_rule1,stemi$SNOMED_STEMI)
kable_confusion_matrix(table_list_snomed[[i]], criteria_name[[i]], stemi_table_name)

i<- i+1
table_list_snomed[[i]] <- table(stemi$journey_acs_rule2,stemi$SNOMED_STEMI)

i<- i+1
# table_list_snomed[[i]] <- table(stemi$journey_above_normal_trop,stemi$SNOMED_STEMI)
# 
# i<- i+1
table_list_snomed[[i]] <- table(stemi$journey_acs_rule1 | stemi$journey_acs_rule2,stemi$SNOMED_STEMI)

i<- i+1
# table_list_snomed[[i]] <- table(stemi$journey_acs_rule1 | stemi$journey_above_normal_trop,stemi$SNOMED_STEMI)
# 
# i<- i+1
table_list_snomed[[i]] <- table(cohort_flags$CRITERIA,cohort_flags$SNOMED_STEMI)

i<- i+1
table_list_snomed[[i]] <- table(new_cohort_flags$CRITERIA_NEW,new_cohort_flags$SNOMED_STEMI)

i<- i+1
table_list_snomed[[i]] <- table(as.logical(new_cohort_flags$CF1_derived),new_cohort_flags$SNOMED_STEMI)

i<- i+1
table_list_snomed[[i]] <- table(as.logical(new_cohort_flags$TROPONIN_FLAG),new_cohort_flags$SNOMED_STEMI)

i<- i+1
table_list_snomed[[i]] <- table(as.logical(new_cohort_flags$CF7_derived),new_cohort_flags$SNOMED_STEMI)

i<- i+1
table_list_snomed[[i]] <- table(as.logical(new_cohort_flags$CF1_derived) | as.logical(new_cohort_flags$TROPONIN_FLAG),new_cohort_flags$SNOMED_STEMI)

i<- i+1
table_list_snomed[[i]] <- table(as.logical(new_cohort_flags$CF1_derived) | as.logical(new_cohort_flags$CF7_derived),new_cohort_flags$SNOMED_STEMI)

i<- i+1
table_list_snomed[[i]] <- table(as.logical(new_cohort_flags$TROPONIN_FLAG) | as.logical(new_cohort_flags$CF7_derived),new_cohort_flags$SNOMED_STEMI)
```


```{r}
df <- data.frame(
                 precision = double(),
                 sensitivity = double())
for(i in 1:length(table_list_snomed)){
  df %<>% dplyr::add_row(
     precision = precision(table_list_snomed[[i]]),
      sensitivity= sensitivity(table_list_snomed[[i]])
  )
}
df %<>% dplyr::mutate(f_score = 2*precision*sensitivity/(precision + sensitivity))
rownames(df) <- unlist(criteria_name, use.names=FALSE)

knitr::kable(df %>% dplyr::mutate_if(is.numeric, format, digits=2), align="r") %>%
  kableExtra::kable_styling()
```
